---
title: "Machine Learning Project"
author: "GI"
date: "3/25/2021"
output: html_document
---

## Analysis and Prediction of Quality of Physical Activity

The purpose of this markdown is to train a model to predict the "classe" variable of a participants activity, while exploring cross validation, out of sample error, and machine learning.  

## Uploading, Exploring, and Cleaning the Dataset

```{r}
# Install necessary library packages and set the seed for analysis
library(caret)
library(rpart)
library(rpart.plot)
set.seed(84)

# Read data into RStudio, replacing blank cells with NA
training <- read.csv('C:/RStudio/newproj/pml-training.csv',na.strings = c('','NA','#DIV/0!'))
testing <- read.csv('C:/RStudio/newproj/pml-testing.csv',na.strings = c('','NA','#DIV/0!'))

# Remove unnecessary data columns (participant name, participant number, timestamps, etc.)
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]

# Split training data set to create a test set to validate model
cvsplit <- createDataPartition(training$classe,p = 0.7, list = FALSE)
cvtraining <- training[cvsplit,]
cvtesting <- training[-cvsplit,]

# Remove columns in which all values are 0
cvtraining <- cvtraining[,colSums(is.na(cvtraining)) == 0]
cvtesting <- cvtesting[,colSums(is.na(cvtesting)) == 0]
```

## Plot Decision Tree

```{r}
# Create decision tree visualization 
dec.tree <- rpart(classe ~ ., data = cvtraining, method = 'class')
rpart.plot(dec.tree)
```

## Train Model to Use for Prediction

The training subset was then put through k-fold cross validation to resample in groups of 10 before creating a model via random forest modeling methodology.

```{r}
# Create a model using the random forest method after cross validating using a 10x K-fold
train.control <- trainControl(method = 'cv',10)
model <- train(classe ~ ., data = cvtraining,method = 'rf', trControl = train.control)
model
```

## Verify Model Efficacy

Since the accuracy for the optimal model (mtry = 2) is 99+%, the model was then used to predict the classe values for the testing subset created from the original training set.  

```{r}
# Use model to predict test subset and check against actual values
mod.pred <- predict(model,cvtesting)
confusionMatrix(cvtesting$classe,mod.pred)

# Calculate the accuracy of the predicting model
accuracy <- postResample(mod.pred,cvtesting$classe)
```

## Conclusion

The predictive model was shown to have an accuracy of `r accuracy[1]` when compared to the testing subset with an out of sample error of `r 1-accuracy[1]`.  


